---
title: "Final Project Code"
author: "Jordan MS Wong"
date: "12/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Setup 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(neonUtilities)

# Load libraries for geospatial analysis and plotting 
library(raster)   # for importing and viewing raster data
library(rgdal)    # for raster metadata
library(gdalUtils)# for raster wrangling
library(ggplot2)  # for plotting
library(dplyr)    # for data wrangling
library(sf)       # for vector data
```

## Data: Plot locations

  Biomass data is available for the entire field site. However, plant diversity data is only available for three of the study plots. My analysis of plant diversity is therefore limited to only three field plots. Within these three field plots are 1m^2 `subplots` from which plant diversity data was collected. I use `full` in variable names to refer to the entire SJER site; otherwise, variables refer only to the three sample plots. 

  NEON provides shapefiles for the centroids, points, and polygons for all of its field sites. These indicate plots and subplots, as well as extent of a collection of plots or entire sites. Here, I import the vector file containing all of the NEON plot locations, then limit the data to only plots of interest, i.e. those that are within the SJER and that I know have plant diversity data. Not all sites have diversity data, and I will only be using diversity data found at the 1m plot level. So I need to find only the 3 locations at SJER that have diversity data, and within those, find only the 1m plots. The locations are also in a different CRS than the biomass data, so I reprojected the biomass raster to match the location data. 
```{r plot location wrangling}
# Limiting the location data to just those we have diversity data for as well

#import plot points
plot_points <- st_read("data/All_NEON_TOS_Plots_V8/All_NEON_TOS_Plot_Points_V8.shp") # all of the plot points at all of the NEON sites

# Limit to only SJER sites with available diversity data
plot_points_SJER <- filter(plot_points, siteID=="SJER", plotID %in% c("SJER_016", "SJER_012", "SJER_022")) # These are the plots we have diversity data for

# Select full SJER site
plot_points_SJER_full <- filter(plot_points, siteID=="SJER")

### extract easting and northing: for fetching biomass data 
# full site

easting_full <- plot_points_SJER_full$easting
northing_full <- plot_points_SJER_full$northing

# limited site
easting  <- plot_points_SJER$easting
northing <- plot_points_SJER$northing



# find just 1m subplots and give them a unique ID 
# and select only the plots that have diversity data
subplot_points_SJER <- plot_points_SJER %>%
  mutate(plottype = nchar(plot_points_SJER$pointID)) %>% 
  filter(plottype == 6) %>% # plot names with 6 characters correspond to 1m plots
  mutate(uniqueID = paste(plotID, pointID)) # unique ID corresponds to both point and plot 

```
Here, I use the `easting_full` and `northing_full` coordinates for the full site to extract biomass within 20m of every plot at the SJER site. 
```{r full site biomass data}
# Biomass from San Joaquin Experimental Range, 2019
# easting and northing must be matched vectors of UTM coordinates
biomass_data_full <- byTileAOP(dpID="DP3.30016.001", site="SJER", year="2019", easting=easting_full, northing=northing_full, buffer=20, check.size = F)

# Need to combine all of this together
biomass_01 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4110000_Biomass.tif")
biomass_02 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4109000_Biomass.tif")
biomass_03 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4107000_Biomass.tif")
biomass_04 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4106000_Biomass.tif")
biomass_05 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_257000_4111000_Biomass.tif")
biomass_06 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_257000_4110000_Biomass.tif")
biomass_07 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_257000_4109000_Biomass.tif")
biomass_08 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_257000_4108000_Biomass.tif")
biomass_09 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_257000_4107000_Biomass.tif")
biomass_10 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4112000_Biomass.tif")

biomass_11 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4111000_Biomass.tif")
biomass_12 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4110000_Biomass.tif")
biomass_13 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4109000_Biomass.tif")

biomass_14 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4108000_Biomass.tif")
biomass_15 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4107000_Biomass.tif")
biomass_16 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_255000_4112000_Biomass.tif")
biomass_17 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_255000_4111000_Biomass.tif")
biomass_18 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_255000_4110000_Biomass.tif")
biomass_19 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_255000_4109000_Biomass.tif")
biomass_20 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_254000_4112000_Biomass.tif")
biomass_21 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_254000_4110000_Biomass.tif")

biomass_full  <- merge(biomass_01, biomass_02, biomass_03,
                       biomass_04, biomass_05, biomass_06,
                       biomass_07, biomass_08, biomass_09,
                       biomass_10, biomass_11, biomass_12,
                       biomass_13, biomass_14, biomass_15,
                       biomass_16, biomass_17, biomass_18,
                       biomass_19, biomass_20, biomass_21)
crs(biomass_full)
biomass_full_df <- as.data.frame(biomass_full, xy=TRUE)


```

I use the `easting` and `northing` from `plot_points_SJER`, which is limited to only the plots with plant diversity data available.
I reprojected the `subplot_points_SJER` vector to match the biomass rasters (proj=UTM)
```{r biomass data just for the plots with diversity data available}
# Fetch biomass data again, but only for those plots with diversity data available
# specified with easting and northing
byTileAOP(dpID="DP3.30016.001", site="SJER", year="2019", easting=easting, northing=northing, buffer=20, check.size = F)

# Import rasters
biomass_22 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4107000_Biomass.tif")
biomass_23 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4106000_Biomass.tif")

# Merge both rasters together
biomass_subplots <- merge(biomass_22, biomass_23)

# Make data frame for plotting
biomass_subplots_df_2 <- as.data.frame(biomass_subplots, xy=TRUE) # because df1 didn't work

# get CRS to reproject the location vector
crs(biomass_subplots) # proj is UTM instead of longlat

# Check subplot_points CRS
crs(subplot_points_SJER) # projection is longlat 

## Reproject subplot points vector to match `biomass_subplots`
subplot_points_SJER_utm <- st_transform(subplot_points_SJER, crs(biomass_subplots))

# Check CRS
crs(subplot_points_SJER_utm)

```


I load in plant presence and percent cover data using NEON utilities, over the three month survey period. I count the number of unique species found in the study plot as distinct `taxonID`. I then add the `uniqueID` column name and a separate column for the date of the data collection. 
```{r PPPC data}
# Getting all of the information in at once using NEONutilities

PPPC <- loadByProduct(dpID = "DP1.10058.001", site = "SJER", startdate = "2019-03", enddate = "2019-03", check.size = F)

# unlist to local environment so we can see all of the data
list2env(PPPC, .GlobalEnv)

# Add unique ID column to 1m2 data frame
div_1m2Data <- div_1m2Data %>%
  mutate(uniqueID = paste(plotID, subplotID))  # will use this for merge later

species_count_1m <- div_1m2Data %>% 
  group_by(uniqueID) %>% # group based on unique ID
  summarize(n = n_distinct(`taxonID`)) %>%
  mutate(plotID = str_split_fixed(uniqueID, " ", n=2)[,1],
         subplotID = str_split_fixed(uniqueID, " ", n=2)[,2])# number of different species 
species_count_1m


```


```{r extract biomass data}
# convert data frame to sf object for extract function
subplot_points_SJER_utm_sf <- st_as_sf(subplot_points_SJER_utm, coords = c("easting", "northing"), crs = crs(subplot_points_SJER_utm))

# Find mean biomass 
mean_biomass <- extract(x = biomass_subplots,
                        y = subplot_points_SJER_utm_sf,
                        buffer = 20,
                        fun = mean, 
                        df = TRUE)
mean_biomass


# attach back to the subplot points frame 
subplot_points_SJER <- mutate(subplot_points_SJER,
                                biomass_mean = mean_biomass$layer)

subplot_points_SJER_df <- as.data.frame(subplot_points_SJER)
```


```{r try to get one data frame}
all_data <- left_join(subplot_points_SJER_df, species_count_1m,
                      by = c('plotID', 'uniqueID'))


all_data

# THIS IS A NOTE FOR MYSELF -- I CAN'T BELIEVE THIS WORKED ON THE FIRST TRY!!!
```

## Analysis
```{r mean biomass full site}
biomass_mean_full <- biomass_full_df %>%
  summarize(biomass_mean = mean(layer, na.rm = TRUE),
            biomass_sd = sd(layer, na.rm = TRUE))
print(biomass_mean_full)

```

```{r anova}
aov_biomass <- aov(biomass_mean ~ plotID, data = all_data)
summary(aov_biomass)

TukeyHSD(aov_biomass)
```

## Figures
```{r figure 1 biomass and subplots}

biomass_map2 <- ggplot(biomass_subplots_df) +
  geom_raster(aes(x = x, y = y, fill = layer)) +
  geom_sf(data = subplot_points_SJER) +
  coord_sf(xlim = c(-119.718, -119.714), ylim = c(37.075, 37.085)) +
  xlab(" ") +
  ylab(" ") +
  theme(legend.title = element_text("Biomass"))

ggplot(biomass_subplots_df_2) +
  geom_raster(aes(x = x, y = y,  fill = layer)) +
  geom_sf(data = subplot_points_SJER_utm)



biomass_map2
```




```{r figure 2 histogram}
biomass_histogram <- ggplot(biomass_total_df) +
  geom_histogram(aes(layer)) +
  xlab(expression(Biomass~g/m^2)) +
  ylab(expression(Count~1m^2~pixels))
biomass_histogram
```

```{r figure 3 barplot}

subplot_species_count <- ggplot(species_count_1m) +
  geom_bar(stat = "identity", aes(x = uniqueID, y = n, color = month, fill = month)) +
  xlab("Unique Subplot ID") +
  ylab("Species Richness") +
  theme(axis.text.x = element_text(angle = 45))

subplot_species_count
```




```{r figure 4}
diversity_biomass <- ggplot(all_data, aes(x = n, y = biomass_mean, color = plotID)) +
  geom_point() +
  geom_smooth(method="lm")
diversity_biomass

```
```
## Figures 

