---
title: "Final Project Code"
author: "Jordan MS Wong"
date: "12/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
- up to you to extend to multiple years 
- put values of biomass tiles from particular plots in context of entire site 
- for example: for biomass data at particular pieces, are they reflective of the whole site? 
- summary statistics? mean of biomass across entire site (histogram)
- y axis of figure 4 - where do they fall in relation to the site mean 
- anova with regard to species count or biomass? do it with the mean biomass and response variable is mean biomass within 20m of subplot 

- incidentally, i hate the "ltd" tag - can we just take it off and write "full" whenever it's the whole dataset, because we're not really looking at full site data anyway
## Setup 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(neonUtilities)

# Load libraries for geospatial analysis and plotting 
library(raster)   # for importing and viewing raster data
library(rgdal)    # for raster metadata
library(gdalUtils)# for raster wrangling
library(ggplot2)  # for plotting
library(dplyr)    # for data wrangling
library(sf)       # for vector data
```

## Data: Plot locations

  Biomass data is available for the entire field site. However, plant diversity data is only available for three of the study plots. My analysis of plant diversity is therefore limited to only three field sites. 

NEED TO FIX THIS: need a few lines where i specify that the easting/northing are just for the 3 plots with diversity data... need the original easting/northing for the biomass_data query

NEON provides shapefiles for the centroids, points, and polygons for all of its field sites. These indicate plots and subplots, as well as extent of a collection of plots or entire sites. Here, I import the vector file containing all of the NEON plot locations, then limit the data to only plots of interest, i.e. those that are within the SJER and that I know have plant diversity data. Not all sites have diversity data, and I will only be using diversity data found at the 1m plot level. So I need to find only the 3 locations at SJER that have diversity data, and within those, find only the 1m plots. The locations are also in a different CRS than the biomass data. 

```{r limiting plot locations}
# Limiting the location data to just those we have diversity data for as well

#import plot points
plot_points <- st_read("data/All_NEON_TOS_Plots_V8/All_NEON_TOS_Plot_Points_V8.shp")

# Select full SJER site
plot_points_SJER_full <- filter(plot_points, siteID=="SJER")

# Limit to only SJER sites with available diversity data
plot_points_SJER_ltd <- filter(plot_points, siteID=="SJER", plotID %in% c("SJER_016", "SJER_012", "SJER_022")) # These are the plots we have diversity data for

### extract easting and northing: for fetching biomass data 
# full site

easting_full <- plot_points_SJER_full$easting
nothing_full <- plot_points_SJER_full$northing

# limited site
easting_ltd  <- plot_points_SJER_ltd$easting
northing_ltd <- plot_points_SJER_ltd$northing



# find just 1m subplots and give them a unique ID 
# and select only the plots that have diversity data
subplot_points_SJER_ltd <- plot_points_SJER_ltd %>%
  mutate(plottype = nchar(plot_points_SJER$pointID)) %>% 
  filter(plottype == 6) %>% # plot names with 6 characters correspond to 1m plots
  mutate(uniqueID = paste(plotID, pointID)) # unique ID corresponds to both point and plot 

```

## I don't think this is still my figure 1??
```{r figure 1}
subplot_points_map <- ggplot(data = subplot_points_SJER_ltd) +
  geom_sf(aes(color = plotID)) 

subplot_points_map
```
***Figure 1.*** Plot locations and sizes at the San Joaquin Experimental Range. 
```{r full site biomass data}
# Biomass from San Joaquin Experimental Range, 2019
# easting and northing must be matched vectors of UTM coordinates
biomass_data <- byTileAOP(dpID="DP3.30016.001", site="SJER", year="2019", easting=easting_full, northing=northing_full, buffer=20)

# Need to combine all of this together
biomass_01 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4110000_Biomass.tif")
biomass_02 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4109000_Biomass.tif")
biomass_03 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4107000_Biomass.tif")
biomass_04 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4106000_Biomass.tif")
biomass_05 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_257000_4111000_Biomass.tif")
biomass_06 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_257000_4110000_Biomass.tif")
biomass_07 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_257000_4109000_Biomass.tif")
biomass_08 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_257000_4108000_Biomass.tif")
biomass_09 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_257000_4107000_Biomass.tif")
biomass_10 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4112000_Biomass.tif")

biomass_11 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4111000_Biomass.tif")
biomass_12 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4110000_Biomass.tif")
biomass_13 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4109000_Biomass.tif")

biomass_14 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4108000_Biomass.tif")
biomass_15 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_256000_4107000_Biomass.tif")
biomass_16 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_255000_4112000_Biomass.tif")
biomass_17 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_255000_4111000_Biomass.tif")
biomass_18 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_255000_4110000_Biomass.tif")
biomass_19 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_255000_4109000_Biomass.tif")
biomass_20 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_254000_4112000_Biomass.tif")
biomass_21 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_254000_4110000_Biomass.tif")

biomass_total <- merge(biomass_01, biomass_02, biomass_03,
                       biomass_04, biomass_05, biomass_06,
                       biomass_07, biomass_08, biomass_09,
                       biomass_10, biomass_11, biomass_12,
                       biomass_13, biomass_14, biomass_15,
                       biomass_16, biomass_17, biomass_18,
                       biomass_19, biomass_20, biomass_21)
crs(biomass_total)
biomass_total_df <- as.data.frame(biomass_total, xy=TRUE)


```

```{r biomass data just for the plots with diversity data available}
biomass_22 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4107000_Biomass.tif")
biomass_23 <- raster("DP3.30016.001/2019/FullSite/D17/2019_SJER_4/L3/Spectrometer/Biomass/NEON_D17_SJER_DP3_258000_4106000_Biomass.tif")
biomass_subplots <- merge(biomass_22, biomass_23)


# Defining new CRS to use with biomass and location data
crs_latlong <- "+proj=longlat +datum=WGS84 +no_defs"



biomass_subplots_latlong <- projectRaster(biomass_subplots, crs=crs_latlong)
crs(biomass_subplots_latlong)

biomass_subplots_df <- as.data.frame(biomass_subplot_latlong, xy=TRUE)

# Check new crs to be sure
crs(subplot_points_SJER_ltd)


```


```{r figure 1 biomass and subplots}

biomass_map2 <- ggplot(biomass_subplots_df) +
  geom_raster(aes(x = x, y = y, fill = layer)) +
  geom_sf(data = subplot_points_SJER) +
  coord_sf(xlim = c(-119.718, -119.714), ylim = c(37.075, 37.085)) +
  xlab(" ") +
  ylab(" ") +
  theme(legend.title = element_text("Biomass"))


biomass_map2
```




```{r figure 2 histogram}
biomass_histogram <- ggplot(biomass_total_df) +
  geom_histogram(aes(layer)) +
  xlab(expression(Biomass~g/m^2)) +
  ylab(expression(Count~1m^2 pixels))
biomass_histogram
```



I still need to add a column that will read only with month sampled (right now it's as 2019-03-10 or something) in `endDate`
I then need to update my graph so that it is colored by endDate as well as separated by plot. 
How does plant diversity change over 3 months? 
```{r PPPC data}
# Getting all of the information in at once using NEONutilities

PPPC <- loadByProduct(dpID = "DP1.10058.001", site = "SJER", startdate = "2019-03", enddate = "2019-05")

# unlist to local environment so we can see all of the data
list2env(PPPC, .GlobalEnv)

# Add unique ID column to 1m2 data frame
div_1m2Data <- div_1m2Data %>%
  mutate(uniqueID = paste(plotID, subplotID)) # will use this for merge later
 
species_count_1m <- div_1m2Data %>%
  group_by(uniqueID) %>% # group based on unique ID
  summarize(n = n_distinct(`taxonID`)) %>%
  mutate(plotID = str_split_fixed(uniqueID, " ", n=2)[,1],
         subplotID = str_split_fixed(uniqueID, " ", n=2)[,2])# number of different species 
species_count_1m


ggplot(species_count_1m) +
  geom_bar(stat = "identity", aes(x = uniqueID, y = n, color = plotID, fill = plotID)) +
  xlab("Unique Subplot ID") +
  ylab("Species Richness") +
  theme(axis.text.x = element_text(angle = 45))
  
```


```{r extract biomass data}
# Find mean biomass 
mean_biomass <- extract(x = biomass_subplots_latlong,
                                  y = subplot_points_SJER_ltd,
                                  buffer = 20,
                                  fun = mean, 
                                  df = TRUE)
mean_biomass

# attach back to the subplot points frame 
subplot_points_SJER <- mutate(subplot_points_SJER,
                                biomass_mean = mean_biomass$layer)

subplot_points_SJER_df <- as.data.frame(subplot_points_SJER)
```


```{r try to get one data frame}
all_data <- left_join(subplot_points_SJER_df, species_count_1m,
                      by = c('plotID', 'uniqueID'))


all_data

# THIS IS A NOTE FOR MYSELF -- I CAN'T BELIEVE THIS WORKED ON THE FIRST TRY!!!
```



```{r figure 4}
diversity_biomass <- ggplot(all_data, aes(x = n, y = biomass_mean, color = plotID)) +
  geom_point() +
  geom_smooth(method="lm")
diversity_biomass

```
```
## Figures 

